### 2020-03-09 CalNotes.tex
Typo missing X,Z
update CLT with only one Poisson problem

Let us first introduce a technical Lemma.
\begin{lemma}\label{lem:CLT} Let $\mathcal L_\epl$ be the generator of the couple $(X^\epl, Z^\epl)^\top$, i.e.,
	\begin{equation}
		\mathcal L_\epl = -\left(\alpha \cdot V'(x) + \frac1\epl p'\left(\frac{x}{\epl}\right)\right)\partial_x + \frac1\delta (x - z) \partial_z  + \sigma \partial^2_{xx}.
	\end{equation} 
	Moreover, let $\rho^\epl$ be the invariant measure of $(X^\epl, Z^\epl)^\top$ and $\Phi^\epl\colon \R^2 \to \R^N$ be the solution of 
	\begin{equation}\label{eq:CLT_Poisson}
		-\mathcal L_\epl \Phi^\epl = \chi -  \E^{\rho^\epl}[\chi],
	\end{equation}
	for $\chi \colon \R^2 \to \R^N$ satisfying $\E^{\rho^\epl}[\Phi^\epl] = 0$. Then, it holds
	\begin{equation}\label{eq:CLT_PoissonIto}
		\frac{1}{T} \int_0^T  \chi(X_t^\epl, Z_t^\epl) \dd t = \E^{\rho^\epl}[\chi(X^\epl, Z^\epl)] - \frac{R^\epl(T)}{T} + \sqrt{2\sigma}\frac{S^\epl(T)}{T},
	\end{equation}
	where
	\begin{equation}\label{eq:CLT_PoissonRemainders}
		R^\epl(T) \defeq \Phi^\epl(X^\epl_T, Z^\epl_T) - \Phi^\epl(X^\epl_0, Z^\epl_0), \quad S^\epl(T) \defeq \int_0^T \partial_x \Phi^\epl(X_t^\epl, Z_t^\epl) \dd W_t.
	\end{equation}
	Moreover, it holds
	\begin{equation}\label{eq:CLT_Equiv}
		2\sigma \E^{\rho^\epl}[\partial_x \Phi^\epl \otimes \partial_x \Phi^\epl] = \E^{\rho^\epl}[\chi \otimes \Phi^\epl + \Phi^\epl \otimes \chi].
	\end{equation}	
\end{lemma}
\begin{proof} The proof of \eqref{eq:CLT_PoissonIto} and \eqref{eq:CLT_PoissonRemainders} is an application of the Itô formula (see e.g. \cite[Remark 6.17]{PaS08}). For \eqref{eq:CLT_Equiv}, it is possible to show that since $\mathcal L_\epl^*\rho^\epl = 0$ it holds 
	\begin{equation}
		\mathcal L_\epl^* (\Phi^\epl \rho) = 2 \sigma \rho^\epl \partial_{xx}^2 \Phi^\epl - \rho^\epl \mathcal L_\epl\Phi^\epl + 2 \sigma \partial_x \Phi^\epl \partial_x \rho^\epl.
	\end{equation}
	Therefore, an integration by parts yields 
	\begin{equation}
	\begin{aligned}
		\E^{\rho^\epl}[\mathcal L_\epl \Phi^\epl \otimes \Phi^\epl] &= 	\int_{\R}\int_{\R} \Phi^\epl \otimes \mathcal L_\epl^*\left(\Phi^\epl \rho^\epl\right) \dd x \dd z \\
		&= -\int_{\R}\int_{\R}\Phi^\epl \otimes \mathcal L_\epl \Phi^\epl \rho^\epl \dd x \dd z- 2\sigma \int_{\R}\int_{\R} \partial_x \Phi^\epl \otimes \partial_x \Phi^\epl \rho^\epl \dd x \dd z\\
		&= -\E^{\rho^\epl}[\Phi^\epl \otimes \mathcal L_\epl \Phi^\epl] - 2\sigma \E^{\rho^\epl}[\partial_x \Phi^\epl \otimes \partial_x \Phi^\epl]. 
	\end{aligned}
	\end{equation}
	Finally, since $\E^{\rho^\epl}[\Phi] = 0$ 
	\begin{equation}
	\begin{aligned}
		2\sigma \E^{\rho^\epl}[\partial_x \Phi^\epl \otimes \partial_x \Phi^\epl] &= -\E^{\rho^\epl}[\mathcal L_\epl \Phi^\epl \otimes \Phi^\epl + \Phi^\epl \otimes \mathcal L_\epl \Phi^\epl]\\
		&= \E^{\rho^\epl}[\chi \otimes \Phi^\epl + \Phi^\epl \otimes \chi],
	\end{aligned}
	\end{equation}
	which is the desired result.
\end{proof}

@@

	\corr{Idea for the first term:} Let us introduce the notation
	J_1^\epl(T) \defeq \sqrt{T}\left( I_1^\epl(T) - \mathcal M_\epl^{-1} \mathfrak p_\epl\right),
	and let us remark that we can rewrite
	\begin{equation} \label{eq:decomposition_first_term}
	J_1^\epl(T) &= \sqrt{T} \left(\frac1T \int_0^T V'(Z^\epl_t) \otimes V'(X^\epl_t) \dd t\right)^{-1}\left(\frac1T \int_0^T \frac{1}{\epl} p' \left ( \frac{X^\epl_t}{\epl} \right ) V'(Z^\epl_t) \dd t - \mathfrak p_\epl \right) \\
	& \quad + \sqrt{T} \left(\left(\frac1T \int_0^T V'(Z^\epl_t) \otimes V'(X^\epl_t) \dd t\right)^{-1} - \mathcal M_\epl^{-1}\right) \mathfrak p_\epl\\
	&\eqdef J_{1,1}^\epl(T) + J_{1,2}^\epl(T).
	Let us denote by $\Phi_1^\epl$ the solution of \eqref{eq:CLT_Poisson} in Lemma \ref{lem:CLT} with right hand side
	\chi_1(x, z) = \frac 1 \epl p' \left ( \frac x \epl \right ) V'(z).
	Then, denoting as $R_1^\epl(T)$ and $S_1^\epl(T)$ the quantities introduced in \eqref{eq:CLT_PoissonRemainders} with $\Phi_1^\epl$ in place of $\Phi^\epl$, we have
	J_{1,1}^\epl(T)= \sqrt{T} \left(\frac1T \int_0^T V'(Z^\epl_t) \otimes V'(X^\epl_t) \dd t\right)^{-1} \left( -\frac{R_1^\epl(T)}{T} + \sqrt{2\sigma}\frac{S_1^\epl(T)}{T} \right).
	Since \corr{$R_1^\epl(T)$ is bounded independently of $\epl$ (is it clear?)}, we first get by the ergodic theorem
	\lim_{T \to \infty} \left(\frac1T \int_0^T V'(Z^\epl_t) \otimes V'(X^\epl_t) \dd t\right)^{-1} \frac{R_1^\epl(T)}{\sqrt{T}} = 0, \quad \text{a.s.}
	Repeating the same reasoning as for $Q^\epl(T)$ and employing the ergodic theorem, we get 
	\lim_{T \to \infty} J_{1,1}^\epl(T) = \lim_{T \to \infty} \sqrt{2\sigma} \left(\frac1T \int_0^T V'(Z^\epl_t) \otimes V'(X^\epl_t) \dd t\right)^{-1} \frac{S^\epl_1(T)}{\sqrt{T}} = \Lambda_1^\epl \sim \mathcal N(0, \Gamma_1^\epl),
	where the covariance is given by due to \eqref{eq:CLT_Equiv}
	\Gamma_1^\epl &= 2\sigma \mathcal M_\epl^{-1}\E^{\rho^\epl}[\partial_x \Phi_1^\epl(X^\epl, Z^\epl) \otimes \partial_x \Phi_1^\epl(X^\epl, Z^\epl)]\mathcal M_\epl^{-\top} \\
	&= \mathcal M_\epl^{-1}\E^{\rho^\epl}[\Phi_1^\epl(X^\epl, Z^\epl) \otimes \chi_1(X^\epl, Z^\epl) + \chi_1(X^\epl, Z^\epl) \otimes \Phi_1^\epl(X^\epl, Z^\epl)]\mathcal M_\epl^{-\top}.
	Let us now consider the term $J_{1,2}^\epl(T)$ and rewrite it as
	J_{1,2}^\epl(T) =  \sqrt{T} \left(\frac1T \int_0^T V'(Z^\epl_t) \otimes V'(X^\epl_t) \dd t\right)^{-1}\left(\mathfrak p_\epl - \frac1T \int_0^T V'(Z^\epl_t) \otimes V'(X^\epl_t) \mathcal M_\epl^{-1} \mathfrak p_\epl  \dd t\right).
	Let now us denote by $\Phi_2^\epl$ the solution of \eqref{eq:CLT_Poisson} in Lemma \ref{lem:CLT} with right hand side
	\chi_2(x, z) = V'(z) \otimes V'(x) \mathcal M_\epl^{-1} \mathfrak p_\epl.
	Applying \eqref{eq:CLT_PoissonIto} and by the definition \eqref{eq:CLT_DefMEps} of $\mathcal M_\epl$, we get
	\frac1T \int_0^T V'(Z^\epl_t) \otimes V'(X^\epl_t) \mathcal M_\epl^{-1} \mathfrak p_\epl \dd t =  \mathfrak p_\epl - \frac{R_2^\epl(T)}{T} + \sqrt{2\sigma}\frac{S_2^\epl(T)}{T},
	where $R_2^\epl(T)$ and $S_2^\epl(T)$ are the quantities introduced in \eqref{eq:CLT_PoissonRemainders} with $\Phi_2^\epl$ in place of $\Phi^\epl$. Therefore, 
	J_{1,2}^\epl(T) =  \sqrt{T} \left(\frac1T \int_0^T V'(Z^\epl_t) \otimes V'(X^\epl_t) \dd t\right)^{-1}\left(\frac{R_2^\epl(T)}{T} - \sqrt{2\sigma}\frac{S_2^\epl(T)}{T}\right).
	Repeating the same reasoning as for $R_1^\epl(T)$ and $S_1^\epl(T)$ and applying the ergodic theorem, we get
	\lim_{T \to \infty} J_{1,2}^\epl(T) = \Lambda_2^\epl \sim \mathcal N(0, \Gamma_2^\epl),
	where the covariance is given by due to \eqref{eq:CLT_Equiv}
		\Gamma_2^\epl &= 2\sigma \mathcal M_\epl^{-1}\E^{\rho^\epl}[\partial_x \Phi_2^\epl(X^\epl, Z^\epl) \otimes \partial_x \Phi_2^\epl(X^\epl, Z^\epl)]\mathcal M_\epl^{-\top} \\
		&= \mathcal M_\epl^{-1}\E^{\rho^\epl}[\Phi_2^\epl(X^\epl, Z^\epl) \otimes \chi_2(X^\epl, Z^\epl) + \chi_2(X^\epl, Z^\epl) \otimes \Phi_2^\epl(X^\epl, Z^\epl)]\mathcal M_\epl^{-\top}.
	\corr{For the covariances $\Gamma_1^\epl$ and $\Gamma_2^\epl$ we don't know how they behave in the limit $\epl \to 0$. It would be nice if they were vanishing, but they involve the solutions $\Phi_1^\epl$ and $\Phi_2^\epl$, which have r.h.s. depending of $\epl$ (especially $\Phi_1$).}

>>>>

Let us first introduce a technical Lemma.
\begin{lemma}\label{lem:CLT} Let $\mathcal L_\epl$ be the generator of the couple $(X^\epl, Z^\epl)^\top$, i.e.,
	\begin{equation}
		\mathcal L_\epl = -\left(\alpha \cdot V'(x) + \frac1\epl p'\left(\frac{x}{\epl}\right)\right)\partial_x + \frac1\delta (x - z) \partial_z  + \sigma \partial^2_{xx}.
	\end{equation} 
	Moreover, let $\rho^\epl$ be the invariant measure of $(X^\epl, Z^\epl)^\top$ and $\Phi^\epl\colon \R^2 \to \R^N$ be the solution of 
	\begin{equation}\label{eq:CLT_Poisson}
		-\mathcal L_\epl \Phi^\epl = \chi^\epl -  \E^{\rho^\epl}[\chi^\epl(X^\epl, Z^\epl)],
	\end{equation}
	for $\chi^\epl \colon \R^2 \to \R^N$ satisfying $\E^{\rho^\epl}[\Phi^\epl(X^\epl, Z^\epl)] = 0$. Then, it holds
	\begin{equation}\label{eq:CLT_PoissonIto}
		\frac{1}{T} \int_0^T  \chi^\epl(X_t^\epl, Z_t^\epl) \dd t = \E^{\rho^\epl}[\chi^\epl(X^\epl, Z^\epl)] - \frac{R^\epl(T)}{T} + \sqrt{2\sigma}\frac{S^\epl(T)}{T},
	\end{equation}
	where
	\begin{equation}\label{eq:CLT_PoissonRemainders}
		R^\epl(T) \defeq \Phi^\epl(X^\epl_T, Z^\epl_T) - \Phi^\epl(X^\epl_0, Z^\epl_0), \quad S^\epl(T) \defeq \int_0^T \partial_x \Phi^\epl(X_t^\epl, Z_t^\epl) \dd W_t.
	\end{equation}
	Moreover, it holds
	\begin{equation}\label{eq:CLT_Equiv}
		2\sigma \E^{\rho^\epl}[\partial_x \Phi^\epl(X^\epl, Z^\epl) \otimes \partial_x \Phi^\epl(X^\epl, Z^\epl)] = \E^{\rho^\epl}
		\begin{aligned}[t]
		[&\chi^\epl(X^\epl, Z^\epl) \otimes \Phi^\epl(X^\epl, Z^\epl) \\
		&+ \Phi^\epl(X^\epl, Z^\epl) \otimes \chi^\epl(X^\epl, Z^\epl)].
		\end{aligned}
	\end{equation}	
\end{lemma}
\begin{proof} The proof of \eqref{eq:CLT_PoissonIto} and \eqref{eq:CLT_PoissonRemainders} is an application of the Itô formula (see e.g. \cite[Remark 6.17]{PaS08}). For \eqref{eq:CLT_Equiv}, it is possible to show that since $\mathcal L_\epl^*\rho^\epl = 0$ it holds 
	\begin{equation}
		\mathcal L_\epl^* (\Phi^\epl \rho^\epl) = 2 \sigma \rho^\epl \partial_{xx}^2 \Phi^\epl - \rho^\epl \mathcal L_\epl\Phi^\epl + 2 \sigma \partial_x \Phi^\epl \partial_x \rho^\epl.
	\end{equation}
	Therefore, an integration by parts yields 
	\begin{equation}
	\begin{aligned}
		\E^{\rho^\epl}[\mathcal L_\epl \Phi^\epl(X^\epl, Z^\epl) \otimes \Phi^\epl(X^\epl, Z^\epl)] &= 	\int_{\R}\int_{\R} \Phi^\epl \otimes \mathcal L_\epl^*\left(\Phi^\epl \rho^\epl\right) \dd x \dd z \\
		&= -\int_{\R}\int_{\R}\Phi^\epl \otimes \mathcal L_\epl \Phi^\epl \rho^\epl \dd x \dd z- 2\sigma \int_{\R}\int_{\R} \partial_x \Phi^\epl \otimes \partial_x \Phi^\epl \rho^\epl \dd x \dd z\\
		&= 
		\begin{aligned}[t]
		&-\E^{\rho^\epl}[\Phi^\epl(X^\epl, Z^\epl) \otimes \mathcal L_\epl \Phi^\epl(X^\epl, Z^\epl)] \\
		&- 2\sigma \E^{\rho^\epl}[\partial_x \Phi^\epl(X^\epl, Z^\epl) \otimes \partial_x \Phi^\epl(X^\epl, Z^\epl)].
		\end{aligned}
	\end{aligned}
	\end{equation}
	Finally, since $\E^{\rho^\epl}[\Phi(X^\epl, Z^\epl)] = 0$ 
	\begin{equation}
	\begin{aligned}
		2\sigma \E^{\rho^\epl}[\partial_x \Phi^\epl(X^\epl, Z^\epl) \otimes \partial_x \Phi^\epl(X^\epl, Z^\epl)] &= -\E^{\rho^\epl}
		\begin{aligned}[t]
		[&\mathcal L_\epl \Phi^\epl(X^\epl, Z^\epl) \otimes \Phi^\epl(X^\epl, Z^\epl) \\
		& + \Phi^\epl(X^\epl, Z^\epl) \otimes \mathcal L_\epl \Phi^\epl(X^\epl, Z^\epl)]
		\end{aligned}
		\\
		&= \E^{\rho^\epl}
		\begin{aligned}[t]
		[&\chi^\epl(X^\epl, Z^\epl) \otimes \Phi^\epl(X^\epl, Z^\epl) \\
		&+ \Phi^\epl(X^\epl, Z^\epl) \otimes \chi^\epl(X^\epl, Z^\epl)],
		\end{aligned}
	\end{aligned}
	\end{equation}
	which is the desired result.
\end{proof}

@@

	\corr{Idea for the first term:} Let us introduce the notation
	\begin{equation}
	J_1^\epl(T) \defeq \sqrt{T}\left( I_1^\epl(T) - \mathcal M_\epl^{-1} \mathfrak p_\epl\right),
	\end{equation}
	and let us remark that we can rewrite
	\begin{equation} \label{eq:decomposition_first_term}
	\begin{aligned}
	J_1^\epl(T) = \sqrt{T} &\left(\frac1T \int_0^T V'(Z^\epl_t) \otimes V'(X^\epl_t) \dd t\right)^{-1}\\
		&\quad \times \left(\frac1T \int_0^T \left( \frac{1}{\epl} p' \left ( \frac{X^\epl_t}{\epl} \right ) V'(Z^\epl_t)	- V'(Z^\epl_t) \otimes V'(X^\epl_t) \mathcal M_\epl^{-1} \mathfrak p_\epl \right) \dd t	 \right).
	\end{aligned}
	\end{equation}
	Let us denote by $\Phi^\epl$ the solution of \eqref{eq:CLT_Poisson} in Lemma \ref{lem:CLT} with right hand side
	\begin{equation}
	\chi^\epl(x, z) = \frac 1 \epl p' \left ( \frac x \epl \right ) V'(z) - V'(z) \otimes V'(x) \mathcal M_\epl^{-1} \mathfrak p_\epl,
	\end{equation}
	and note that $\E^{\rho^\epl}[\chi^\epl(X^\epl,Z^\epl)] = 0$. Then we have
	\begin{equation}
	J_1^\epl(T)= \sqrt{T} \left(\frac1T \int_0^T V'(Z^\epl_t) \otimes V'(X^\epl_t) \dd t\right)^{-1} \left( -\frac{R^\epl(T)}{T} + \sqrt{2\sigma}\frac{S^\epl(T)}{T} \right),
	\end{equation}
	where $R^\epl(T)$ and $S^\epl(T)$ are defined in \eqref{eq:CLT_PoissonRemainders}. Since \corr{$R^\epl(T)$ is bounded independently of $\epl$ (is it clear?)}, we first get by the ergodic theorem
	\begin{equation}
	\lim_{T \to \infty} \left(\frac1T \int_0^T V'(Z^\epl_t) \otimes V'(X^\epl_t) \dd t\right)^{-1} \frac{R^\epl(T)}{\sqrt{T}} = 0, \quad \text{a.s.}
	\end{equation}
	Repeating the same reasoning as for $Q^\epl(T)$ and employing the ergodic theorem and the Slutsky's theorem, we get 
	\begin{equation}
	\lim_{T \to \infty} J_1^\epl(T) = \lim_{T \to \infty} \sqrt{2\sigma} \left(\frac1T \int_0^T V'(Z^\epl_t) \otimes V'(X^\epl_t) \dd t\right)^{-1} \frac{S^\epl(T)}{\sqrt{T}} = \Lambda_1^\epl \sim \mathcal N(0, \Gamma_1^\epl),
	\end{equation}
	where the covariance is given by due to \eqref{eq:CLT_Equiv}
	\begin{equation}
	\begin{aligned}
	\Gamma_1^\epl &= 2\sigma \mathcal M_\epl^{-1}\E^{\rho^\epl}[\partial_x \Phi^\epl(X^\epl, Z^\epl) \otimes \partial_x \Phi^\epl(X^\epl, Z^\epl)]\mathcal M_\epl^{-\top} \\
	&= \mathcal M_\epl^{-1}\E^{\rho^\epl}[\Phi^\epl(X^\epl, Z^\epl) \otimes \chi^\epl(X^\epl, Z^\epl) + \chi^\epl(X^\epl, Z^\epl) \otimes \Phi^\epl(X^\epl, Z^\epl)]\mathcal M_\epl^{-\top}.
	\end{aligned}
	\end{equation}
	\corr{For the covariance $\Gamma_1^\epl$ we don't know how it behaves in the limit $\epl \to 0$. It would be nice if it was vanishing, but it involves the solution $\Phi^\epl$, which has r.h.s. depending of $\epl$. We noticed that the homogenized problem of
	\begin{equation}
		-\mathcal L_\epl \Phi^\epl = \chi^\epl,
	\end{equation}
	is given by
	\begin{equation}
		-\mathcal L_0 \Phi^0 = 0,
	\end{equation}
	whose zero-mean solution is $\Phi^0 = 0$.}
